---
output:
  html_document:
    df_print: paged
---

Load the required packages
```{r, warning = FALSE, error = FALSE, message = FALSE}
library(GeoLight)
library(TwGeos)
library(FLightR)
library(MASS)  #needed for fitting distributions
library(maptools)
library(MASS)
library(raster)
library(sp)
library(SGAT)
library(LLmig)
```

Read in the data files & create vector of 'birdnames' using the file names.
```{r}
lightfiles <- list.files("Data",full.names = TRUE)
birdnames <- list.files("Data")
```

Create a variable that store the number of birds and a list to store the raw light data
```{r}
nBirds <- length(lightfiles)
lightData <- vector('list',nBirds)
```

Create an object that stores the location of birds that have .lig and .lux data types
```{r}
ligBirds <- grep(lightfiles, pattern = ".lig")
luxBirds <- grep(lightfiles, pattern = ".lux")
```

### Read in the raw light-level data for all birds
```{r}
lightData[ligBirds] <- lapply(lightfiles[ligBirds],readLig)
lightData[luxBirds] <- lapply(lightfiles[luxBirds],readMTlux)
```

Save only the columns named Date and Light
```{r}
lightData[ligBirds] <- lapply(lightData[ligBirds],
                              FUN = function(x){subset(x,select=c("Date","Light"))}) 

# Have look
str(lightData,1)
```

### Capture locations 
```{r}
CapLocs <- array(NA,c(nBirds,2))
rownames(CapLocs)<-birdnames

CapLocs[1,] <- cbind(-149.804,  61.158)
CapLocs[2,] <- cbind(-149.804,  61.158)
CapLocs[3,] <- cbind(-149.771,  61.089)
CapLocs[4,] <- cbind(-149.771,  61.089)
CapLocs[5,] <- cbind(-149.804,  61.158)
CapLocs[6,] <- cbind(-149.804,  61.158)
CapLocs[7,] <- cbind(-149.804,  61.158)
CapLocs[8,] <- cbind(-149.750,  61.284)
CapLocs[9,] <- cbind(-149.750,  61.284)
CapLocs[10,] <- cbind(-149.750,  61.284)
CapLocs[11,] <- cbind(-149.751,  61.2837)
```

### Set the seed
```{r}
# The time when we know the geolocator recorded no light.
lapply(lightData,head)

# Find dates for the see
seed <- lapply(lightData, FUN = function(x){as.POSIXct(c(paste0(format(x[1,1],"%Y"),"-12-01 04:00:00"),
                                         paste0(format(x[nrow(x),1],"%Y"),"-01-01 04:00:00"),
                                        paste0(format(x[nrow(x),1],"%Y"),"-02-01 04:00:00")),tz = "GMT")})
```

### Assign twilights
```{r}
# Assign twilights to all birds at once
twl <- mapply(lightData,seed, SIMPLIFY = FALSE, FUN = function(x,y){findTwilights(tagdata = x, 
                                                                threshold = 1.5, 
                                                                include = y,
                                                                dark.min = 0)})

```

### Edit transitions
```{r}
#Moves or deletes outliers
twl <- lapply(twl, twilightEdit, 
                    window = 4,           # two days before and two days after
                    outlier.mins = 45,    # difference in mins
                    stationary.mins = 25, # are the other surrounding twilights within 25 mins of one another
                    plot = FALSE)

#In order to take advantage of the edits - you can subset the data using when Deleted == FALSE this removes 
#any of the twilights which were deleted.
#Here we adjust the twilight time back to account for the max light level readings every # minutes
twl <- lapply(twl,twilightAdjust, interval = 300) # The unit here is seconds

# remove deleted
twl <- lapply(twl,subset,Deleted == FALSE)
```

## Zenith angles
```{r}
# These dates should be changed a bit - as of right now the whole year is used as calibration dates
# what you want are dates when you know that the bird is at the deployment site. This can be a bit 
# tricky for near arctic breeding birds when tags are deployed during the summer. I moved some things
# around. Here, I use the lightImage and the cal.data to determine when the bird arrives in the spring
# I use those data until incubation starts to calculate zenith angles. Just after deployment doesn't 
# work that well in this case because of the nearly constant light issues. 

# Create a vector of the dates the geolocator was recording data
tm <- seq(lightData[[11]][1,1], lightData[[11]][nrow(lightData[[11]]),1], by = "day")
head (tm)

# Create a vector of TRUE /FALSE to determine whether it's sunrise or sunset
rise <- rep(c(TRUE, FALSE), length(tm))

# making predicted twilight times given location and zenith #
cal.data<- data.frame(Twilight = twilight(rep(tm, each = 2),
                                          lon = CapLocs[11,1], 
                                          lat = CapLocs[11,2], 
                                          rise = rise, zenith = 95),
                      Rise = rise) 
```

Calibration dates
```{r}
calib.dates <- as.POSIXct(c("2017-05-11","2017-05-25"), tz = "GMT")

# Extract twilight data during calibration period
cal.data<-subset(twl[[11]],twl[[11]]$Twilight>=calib.dates[1] & twl[[11]]$Twilight<=calib.dates[2])

# Determine the sun elevation angle - here called the Zenith angle #
# Calculate solar time from calibration data 
sun  <- solar(cal.data[,1])

# Adjust the solar zenith angle for atmospheric refraction
z<- refracted(zenith(sun = sun,
                     lon = CapLocs[11,1],
                     lat = CapLocs[11,2]))

twl_t <- twilight(cal.data[,1], CapLocs[11,1], CapLocs[11,2], rise = cal.data[,2], zenith = quantile(z,probs=0.5))

# Determine the difference in minutes from when the sun rose and the geolocator said it rose 
twl_dev <- ifelse(cal.data$Rise, as.numeric(difftime(cal.data[,1], twl_t, units = "mins")),
                  as.numeric(difftime(twl_t, cal.data[,1], units = "mins")))

# Describe the distribution of the error 
fitml <- fitdistr(abs(twl_dev), "Log-normal")
head (fitml)

# save the Twilight model parameters
alpha<- c(fitml$estimate[1], fitml$estimate[2]) 
head (alpha)

# Make some plots to visualize the data 
seq <- seq(0,60, length = 100)
hist(twl_dev, freq = F,
     yaxt="n",
     ylim = c(0, 0.15),
     xlim = c(0, 60),
     breaks=20,
     main = "",
     xlab = "Twilight error (mins)")
axis(2,las=2)
lines(seq, dlnorm(seq, alpha[1], alpha[2]), col = "firebrick", lwd = 3,lty = 2)
```

```{r}
# Store the zenith (sun-elevation angle)
zenith0 <-quantile(z,prob=0.5)
zenith1 <-quantile(z,prob=0.95)
```

```{r}
tol <- array(NA,c(nBirds,2))
rownames(tol)<-birdnames
tol[1,] <- cbind(0.14,0.20)
tol[2,] <- cbind(0.16,0.09)
tol[3,] <- cbind(0.25,0.16)
tol[4,] <- cbind(0.16,0.09)
tol[5,] <- cbind(0.16,0.21)
tol[6,] <- cbind(0.20,0.20)
tol[7,] <- cbind(0.16,0.20)
tol[8,] <- cbind(0.22,0.2)
tol[9,] <- cbind(0.16,0.09)
tol[10,] <- cbind(0.07,0.02)
tol[11,] <- cbind(0.16,0.09)
```

Initial path
```{r}
path <- mapply(twl,tol[,1],tol[,2], SIMPLIFY = FALSE,
               FUN = function(x,y,z){thresholdPath(twilight = x$Twilight, 
                      rise = x$Rise, 
                      zenith=zenith0,
                      tol= c(y,z))})
```

```{r}
# Set what xlim and ylim values you need to span the range of your dataset
xlim <- c(-180, -30)
ylim <- c(-80, 90)

# Land Mask-------------------------------------------------------------------------------------------------------

land.mask <- function(shape, xlim, ylim, res = c(0.25,0.25), land = TRUE) {
  #Create a raster grid of "NA" that fills the x and y limits. Use a lat/lon projections
  #
  # Changing the setting the resolution is a better approach than 1/n degrees. 
  # This ensures that the resolution is square - that's not always the case with 1/n
  r <- raster(xmn = xlim[1], 
              xmx = xlim[2], 
              ymn = ylim[1],
              ymx = ylim[2], 
              res = res, 
              crs = proj4string(shape))

  #Replace the NA values with 1 where there is land
  #This code is kind of complicated because it allows for grids that wrap around the date line

  r <- cover(rasterize(shape, r, 1, silent = TRUE), 
             rasterize(shape, r, 1, silent = TRUE), 
             rasterize(shape, r, 1, silent = TRUE))
  
  #make the raster a matrix with column order reversed and NA set to TRUE
  r <- as.matrix(is.na(r))[nrow(r):1, ]
  
  if (land) #reverse the TRUE/FALSE if land is set to TRUE
    r <- !r

xbin <- seq(xlim[1], xlim[2], length = ncol(r) + 1)
ybin <- seq(ylim[1], ylim[2], length = nrow(r) + 1)
  
  function(p) {
    r[cbind(.bincode(p[, 2], ybin), .bincode(p[, 1], xbin))]
  }
}

data(wrld_simpl,package = "maptools")

# Remove Caribbean from potential stop-over locations #
NS_AM <- wrld_simpl[which(wrld_simpl$SUBREGION == 13 | 
                          wrld_simpl$SUBREGION == 21 |
                          wrld_simpl$SUBREGION == 5),]


is.dist <- land.mask(shape=NS_AM,
                     xlim = xlim,
                     ylim = ylim,
                     land = TRUE)

log.prior <- function(p) {
  f <- is.dist(p)
  ifelse(f | is.na(f), 0, -10)
}
```

```{r}
fit <- vector('list',nBirds)

beta <- c(0.7, 0.08)

for(i in 1:nBirds){
model<- thresholdModel(twl[[i]]$Twilight,
                       twl[[i]]$Rise,
                       twilight.model="ModifiedLogNormal",
                       alpha=alpha,
                       beta=beta,
                       # Here is where we set the constraints for land
                       logp.x = log.prior, 
                       logp.z = log.prior,
                       x0=path[[i]]$x,
                       z0=trackMidpts(path[[i]]$x),
                       zenith=zenith0)


# This defines the error distribution around each location 

proposal.x <- mvnorm(S=diag(c(0.0025,0.0025)),n=nlocation(path[[i]]$x))
proposal.z <- mvnorm(S=diag(c(0.0025,0.0025)),n=nlocation(trackMidpts(path[[i]]$x)))          

fit[[i]] <- estelleMetropolis(model = model,
                         proposal.x = proposal.x,
                         proposal.z = proposal.z,
                         iters=2000, # This value sets the number of iterations to run
                         thin=5,
                         chains=3)

xsum <- locationSummary(fit[[i]]$x,collapse = TRUE)
zsum <- locationSummary(fit[[i]]$z,collapse = TRUE)

x0 = cbind(xsum$'Lon.50%',xsum$'Lat.50%')
z0 = cbind(zsum$'Lon.50%',zsum$'Lat.50%')

fit[[i]] <- estelleMetropolis(model = model,
                         proposal.x = proposal.x,
                         proposal.z = proposal.z,
                         x0 = x0,
                         z0 = z0,
                         iters=2000, # This value sets the number of iterations to run
                         thin=5,
                         chains=3)

#Final Run

xsum<-locationSummary(fit[[i]]$x,collapse = TRUE)
zsum<-locationSummary(fit[[i]]$z,collapse = TRUE)

proposal.x <- mvnorm(chainCov(fit[[i]]$x),s=0.1)
proposal.z <- mvnorm(chainCov(fit[[i]]$z),s=0.1)

fit[[i]] <- estelleMetropolis(model = model,
                         proposal.x = proposal.x,
                         proposal.z = proposal.z,
                         x0 = x0,
                         z0 = z0,
                         iters=5000, # This value sets the number of iterations to run
                         thin=1,
                         chains=3)
}
```